% Copyright 2024 Kieran W Harvie. All rights reserved.

\section{Bézier Curves}
Given $n+1$ control points $\mathbf{P}_k$ the $n$th order Bézier curve is given by:
\[
	f(t) = \sum_{k=0}^nb_{k,n}(t)\mathbf{P}_k,\quad t\in [0,1]
\]
Note that sometimes the function itself,
without the domain,
is also called a Bézier curve.
\\

The lower order Bézier curves have their expected special names,
for example a Linear Bézier curve is:
\[f(t)=(1-t)\mathbf{P}_0+t\mathbf{P}_1\]
A Quadratic Bézier curve is:
\[f(t)=(1-t)^2\mathbf{P}_0+2t(1-t)\mathbf{P}_1+t^2\mathbf{P}_2\]
And a Cubic Bézier curve is:
\[f(t)=(1-t)^3\mathbf{P}_0+3t(1-t)^2\mathbf{P}_1+3t^2(1-t)\mathbf{P}_2+t^3\mathbf{P}_3\]
These curves are the most common ones you will see in practice,
Cubic Bézier curves in particular,
as it has wide support across many programs.

\begin{center}
\begin{tikzpicture}
	\draw[red] (0,0) .. controls (1,3) and (4,0) .. (5,3);
	\filldraw (0,0) node[above]{} circle(2pt);
	\filldraw (1,3) node[above]{} circle(2pt);
	\filldraw (4,0) node[above]{} circle(2pt);
	\filldraw (5,3) node[above]{} circle(2pt);
\end{tikzpicture}

An example Cubic Bézier curve with control points shown.
\end{center}

\subsubsection{Interpreting the control points}
There's two interpretations on the control points of a Bézier curve.
\\

First is that the points determine the location and derivatives of the start and end points.
The curve starts at $\mathbf{P}_0$ and ends at $\mathbf{P}_n$.
Its derivative at the start lies on $\mathbf{P}_1-\mathbf{P}_0$ and at the end on $\mathbf{P}_{n-1}-\mathbf{P}_n$.
Its second derivative at the start is controlled by $\mathbf{P}_2$ and end by $\mathbf{P}_{n-2}$, etc.
A corollary of this is that two curves who share all their control points less than $k$ have the same $k$ or less derivatives at the start of the curves,
likewise for $k$ and above and $n-k$ and the curves end.
\\

Second is that the points pull the curve towards them and with a strength and variability determined by $b_{k,n}$.
To see this let $f$ and $\hat{f}$ be $n$ Bézier Curves that differ only with control point $\mathbf{P}_k$ and $\hat{\mathbf{P}_k}$, then we have.
\[
	\hat{f}(t) = b_{k,n}(t)(\hat{\mathbf{P}}_k-\mathbf{P}_k)+f(t)
\]
Since $b_{k,n}$ is positive the point $\hat{f}(t)$ is always the point $f(t)$ displaced in the direction of $\hat{\mathbf{P}_k}-\mathbf{P}_k$ with the maximum displacement happening at $\frac{k}{n}$.

\subsubsection{Convex Hull}

\subsubsection{Affine Transform}
Let $T$ be some affine transform then $b_{k,n}$ being a partition of unity mean:
\[\begin{aligned}
	T(f(t)) =& T\left(\sum_{k=0}^nb_{k,n}(t)\mathbf{P}_k\right)\\
	=& \sum_{k=0}^nb_{k,n}(t)T(\mathbf{P}_k)\\
\end{aligned}\]
In other words,
given some control points the affine transform a Bézier curve is the Bézier curves of the transformed control points.

\subsubsection{Subdivision and Extension:}
By applying the inner product property of Bernstein basis polynomials we obtain:
\[\begin{aligned}
	f(\tau t) =& \sum_{k=0}^n\mathbf{P}_kb_{k,n}(\tau t)\\
	=&\sum_{k=0}^n\mathbf{P}_k\sum_{l=k}^nb_{l,n}(t)b_{k,l}(\tau)\\
	=&\sum_{l=0}^nb_{l,n}(t)\sum_{k=0}^l\mathbf{P}_kb_{k,l}(\tau)\\
\end{aligned}\]
And similarly for the reflected inner product property:
\[\begin{aligned}
	f(\tau +(1-\tau)t)=& \sum_{k=0}^n\mathbf{P}_kb_{k,n}(\tau+(1-\tau)t)\\
	=&\sum_{k=0}^l\mathbf{P}_k\sum_{l=0}^kb_{k-l,n-l}(\tau)b_{l,n}(t)\\
	=&\sum_{l=0}^nb_{l,n}(t)\sum_{k=l}^n\mathbf{P}_kb_{k-l,n-l}(\tau)\\
\end{aligned}\]
When $\tau\in[0,1]$ the functions $f(\tau t)$ and $f(\tau + (1-\tau)t)$ can be interpreted as subdividing $f(t)$ into two function.
Once which matches the values of $f$ from $0$ to $\tau$ and one which matches from $\tau$ to $1$.
We can use the inner sum to extract the control points for each subdivision,
explicitly in the cubic case:
\[\begin{array}{|c|c|}
	\hline
	f(t)&f(\tau t)\\
	\hline
	\mathbf{P}_3&(1-\tau)^3\mathbf{P}_3+3\tau(1-\tau)^2\mathbf{P}_2+3\tau^2(1-\tau)\mathbf{P}_1+\tau^3\mathbf{P}_0\\
	\mathbf{P}_2&(1-\tau)^2\mathbf{P}_2+2\tau(1-\tau)\mathbf{P}_1+\tau^2\mathbf{P}_0\\
	\mathbf{P}_1&(1-\tau)\mathbf{P}_1+\tau \mathbf{P}_0\\
	\mathbf{P}_0&\mathbf{P}_0\\
	\hline
\end{array}\]
\[\begin{array}{|c|c|}
	\hline
	f(t)&f(\tau+(1-\tau)t)\\
	\hline
	\mathbf{P}_3&\mathbf{P}_3\\
	\mathbf{P}_2&\tau \mathbf{P}_3+(1-\tau)\mathbf{P}_2\\
	\mathbf{P}_1&\tau^2\mathbf{P}_3+2\tau(1-\tau)\mathbf{P}_2+(1-\tau)^2\mathbf{P}_1\\
	\mathbf{P}_0&\tau^3\mathbf{P}_3+3\tau^2(1-\tau)\mathbf{P}_2+3\tau(1-\tau)^2\mathbf{P}_1+(1-\tau)^3\mathbf{P}_0\\
	\hline
\end{array}\]
When $\tau$ is outside $[0,1]$ the functions can be interpreted extending $f(t)$ to between $0$ and $\tau$.
Note that care must be taken to the sign of $\tau$ as $[-2,0]$ and $[0,2]$ are different as the former extends from the $0$ end and the later from the $1$ end.

\subsubsection{Useful Curves:}
In the Bernstein basis polynomials section we calculated some sums that are useful starting points and sanity checks when writing algorithms.
These sums are similarly useful for calculating Bézier curves.
For example,
consider control a collection of regularly spaced control points:
\[\mathbf{P}_k = \mathbf{A}+k\mathbf{B}\]
Then Bézier curve is given by:
\[\begin{aligned}
	f(t) &= \sum_{k=0}^nb_{k,n}(t)\mathbf{P}_k\\
	&= \sum_{k=0}^nb_{k,n}(t)(\mathbf{A}+t\mathbf{B})\\
	&= \mathbf{A}\sum_{k=0}^nb_{k,n}(t)+\mathbf{B}\sum_{k=0}^nkb_{k,n}(t)\\
	&= \mathbf{A}+nt\mathbf{B}\\
\end{aligned}\]
