% Copyright 2024 Kieran W Harvie. All rights reserved.

\section{Introduction}
Bézier curves and surfaces are an elegant way to enhance our models and renders with natural curves.
The foremost resource on the application of curves and surfaces to these problems,
the resource that many an intrigued person is pointed to,
is "Curves and Surfaces for CAGD"\cite{CAGD}.
This work is phenomenal.
Providing a wonderful treatment on not only Bézier curves and surfaces but surrounding theory of blossoms,
the algorithm poetically call "the blossoming",
and along with other more general curves and surfaces.
I can't understate how much anyone reading this document would find value in reading it too.
\\

However,
while reading the book,
I was continuously compelled to rewrite the presented mathematics in a style more familiar to me,
this document is the end result of that compulsion.
This style,
along with targeted level of audience knowledge,
is the esteemed undergraduate textbook "Algebra" by Artin\cite{artin}.
The results needed from Artin are nothing more complex than knowing what an $R$-module is or why every vector space has basis.
Additionally any background material outside of Artin has been included in the appendix,
A choice Artin made himself.
\\

This book will begin by describing various Bézier curves and surfaces along with the properties Bernstein basis polynomials
under the general umbrella of `Bézier Geometry'.
The we will define Blossoms and the blossoming.
Finally we will combine the two and apply blossoms too Bézier Geometry.
