% Copyright 2023 Kieran W Harvie. All rights reserved.

\section{Introduction}
This document is the main place I collect notes on mathematic and mathematic adjacent topics I felt the desire to write down.
Either because I want to come back to the notes later,
felt I needed some revision on the covered topic,
or was interested in writing the notes for their own sakes.
\\

Because of the disparate topics and reasons for writing the notes there is little structure shared across the whole document.
The only real structure is that sections are grouped into years and organized by most recent to oldest,
based on when they are first written,
however older sections may still be modified without updating this order.
\\

I've removed the table of contents from within the document itself because it the unstructured nature of the document made it difficult to navigate.
But the structure is still present in the file and you are able to use the `Table of Contents` or `Document Outline' feature of your PDF viewer. 

\subsection{Showcase}
I'm particularly fond of and wish to showcase the following section to onlookers:
\begin{itemize}
	\item \hyperref[showcase:magnetism]{\bf Magnetism} 
		Particularly nice mix of text, diagrams, and equations.
	\item \hyperref[showcase:rational_residue]{\bf Rational Residue} 
		Made over three commits each separated by multiple days. 
		Each subsequent commit shows a growth in understanding over the previous,
		going from `neat trick' to `proper method' to 'generalization'.
	\item \hyperref[showcase:sqrt_limit]{\bf Square Root Limit} 
		An interesting result that shows three solutions to the problem with different ratios of algebra to analysis.
		%orthogonal group
\end{itemize}
